Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-11-07 manuscript/manuscript.Rmd

Despite the weak climate effects, including climate covariates did increase the \textcolor{blue}{mean} accuracy of forecasts for two species: *B. gracilis* and *Poa secunda* (\textcolor{blue}{Fig. S3}).
\textcolor{blue}{However, skill increases were only statistically significant at $\alpha$ = 0.05 for the QBM for \emph{B. gracilis} QBM ($t_{(279)}$ = 1.80, \emph{P} = 0.037).}
Forecast error decreased significantly with the inclusion of climate covariates for the *B. gracilis* IPM ($t_{(280)}$ = -3.72, *P* = \textcolor{blue}{0.01}) and QBM ($t_{(280)}$ = -3.34, *P* < 0.0001), and for the *Poa secunda* IPM ($t_{(196)}$ = -1.90, *P* \textcolor{blue}{= 0.015}) and QBM ($t_{(196)}$ = -2.47, *P* = 0.007) (Fig. S1, Supporting Information).
In no case did including climate covariates significantly decrease forecast skill (Table S21), despite small changes in the mean skill (Fig. 2, \textcolor{blue}{Fig. S3}).

@@

With climate covariates included and using mean parameter values, the accuracy of both models' forecasts declined as the distance between the last observation and the forecast increased, but they did so at similar rates (Fig. 4).
The only exception is for *Poa secunda*, where QBM forecast accuracy remained steady as the temporal distance of the forecast increased, whereas IPM forecast accuracy declined (Fig. 4).
The forecast horizons were short: forecast accuracy fell below $\rho$ = 0.5 after one year for the IPM for most species, and after four years, at most, for the QBM (Fig. 4).
Across the different temporal distances from the observation to the forecast, the IPM was never more accurate than the QBM (\emph{P} > 0.05 for all one-sided \emph{t}-tests, Table S23).
\textcolor{blue}{Likewise, the QBM was rarely more accurate the IPM, the only exceptions being for \emph{H. comata} at a temporal distance of three years ($t_{(98)}$ = 2.04, \emph{P} = 0.032) and \emph{B. gracilis} at a temporal distance of eight years ($t_{(37)}$ = 1.67, \emph{P} = 0.05) (Table S24).}
\st{There were some cases where the QBM was more accurate than the IPM for \emph{Poa secunda}, but neither model exceeded the forecast proficiency threshold by a large margin (Fig. 4, Table S24).}

@@

Acknowledgments
---------------
This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370), award MSB-1241856 to MBH, and a CAREER award to PBA (DEB-1054040). We thank the original mappers of the permanent quadrats in Montana and the digitizers in the Adler lab, without whom this work would not have been possible. Informal conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a series of meetings between the Adler and Weecology labs at USU sharpened our thinking. Brittany Teller provided comments that improved our manuscript. Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged. Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S. government. This research was supported by the Utah Agricultural Experiment Station, Utah State University, and approved as journal paper number 8917. \textcolor{blue}{And, thanks to Andy Kleinhesselink for discovering an error in our computer code before this paper was published.}

>>>>

Despite the weak climate effects, including climate covariates did increase the \textcolor{blue}{mean} accuracy of forecasts for two species: *B. gracilis* and *Poa secunda* (Fig.2 and \textcolor{blue}{Fig. S3, Supporting Information}).
\textcolor{blue}{However, skill increases were only statistically significant at $\alpha$ = 0.05 for the \emph{B. gracilis} QBM ($t_{(279)}$ = 1.80, \emph{P} = 0.037).}
Forecast error decreased significantly with the inclusion of climate covariates for the *B. gracilis* IPM ($t_{(280)}$ = -3.72, *P* = \textcolor{blue}{0.01}) and QBM ($t_{(280)}$ = -3.34, *P* < 0.0001), and for the *Poa secunda* IPM ($t_{(196)}$ = -1.90, *P* \textcolor{blue}{= 0.015}) and QBM ($t_{(196)}$ = -2.47, *P* = 0.007) (Fig. S1, Supporting Information).
In no case did including climate covariates significantly decrease forecast skill (Table S21), despite small changes in the mean skill (Fig. 2, \textcolor{blue}{Fig. S3}).

@@

With climate covariates included and using mean parameter values, the accuracy of both models' forecasts declined as the distance between the last observation and the forecast increased, but they did so at similar rates (Fig. 4).
The only exception is for *Poa secunda*, where QBM forecast accuracy remained steady as the temporal distance of the forecast increased, whereas IPM forecast accuracy declined (Fig. 4).
The forecast horizons were short: forecast accuracy fell below $\rho$ = 0.5 after one year for the IPM for most species, and after four years, at most, for the QBM (Fig. 4).
Across the different temporal distances from the observation to the forecast, the IPM was never more accurate than the QBM (\emph{P} > 0.05 for all one-sided \emph{t}-tests, Table S23).
\textcolor{blue}{Likewise, the QBM was rarely more accurate than the IPM, the only exceptions being for \emph{H. comata} at a temporal distance of three years ($t_{(98)}$ = 2.04, \emph{P} = 0.032) and \emph{B. gracilis} at a temporal distance of eight years ($t_{(37)}$ = 1.67, \emph{P} = 0.05) (Table S24).}
\st{There were some cases where the QBM was more accurate than the IPM for \emph{Poa secunda}, but neither model exceeded the forecast proficiency threshold by a large margin (Fig. 4, Table S24).}

@@

Acknowledgments
---------------
This work was funded by the National Science Foundation through a Postdoctoral Research Fellowship in Biology to ATT (DBI-1400370), award MSB-1241856 to MBH, and a CAREER award to PBA (DEB-1054040). We thank the original mappers of the permanent quadrats in Montana and the digitizers in the Adler lab, without whom this work would not have been possible. Informal conversations with Stephen Ellner, Giles Hooker, Robin Snyder, and a series of meetings between the Adler and Weecology labs at USU sharpened our thinking. Brittany Teller provided comments that improved our manuscript. Compute, storage and other resources from the Division of Research Computing in the Office of Research and Graduate Studies at Utah State University are gratefully acknowledged. Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S. government. This research was supported by the Utah Agricultural Experiment Station, Utah State University, and approved as journal paper number 8917.
